---
title: "01 Data Load"
author: J Andres Gannon, Erik Gartzke, and Jon Lindsay, Center for Peace and Security Studies (cPASS)
date: "`r Sys.Date()`"
output:
  html_notebook:
    fig_height: 8
    fig_width: 12
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
editor_options:
  chunk_output_type: inline
---
<style>
    body .main-container {
        max-width: 100%;
    }
</style>

This is the entry point for the paper "After Deterrence: Explaining Conflict Short of War" 

# Load packages
Pipe operators don't play nice and must be manually loaded up front
```{r}
library(magrittr)
```

# Load data
We load the data from the google spreadsheet tab that has aggregated the info from previous versions. Each row is now a gray zone case and the columns are coded covariates. The distance data represented here was calculated using the code in 02_Distance.Rmd and then inputed into the google spreadsheet.
```{r}
grayzone <- googlesheets::gs_title("Russian wars")
grayzone <- grayzone %>% googlesheets::gs_read("new_cases")
```
# Save file
```{r}
saveRDS(grayzone, paste0(here::here(), "/data/","grayzone_cpass.rds"))
```
