---
title: "03c_REI"
author: J Andres Gannon, Erik Gartzke, and Jon Lindsay, Center for Peace and Security
  Studies (cPASS)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  html_notebook:
    fig_height: 8
    fig_width: 12
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document will look at coverage of Russian electoral inference (1991-2017) from Casey and Ahmad (2017) and identify variables of interest. This is the third of 3 prior datasets used to determine the universe of cases.

The cyber data comes from Casey, Adam; Way, Lucan Ahmad, 2017, "Russian Electoral Interventions, 1991-2017", https://doi.org/10.5683/SP/BYRQQS, Scholars Portal Dataverse

# Preparation
## Load packages
Pipe operators have trouble loading for individual commands
```{r}
library(magrittr)
library(ggplot2)
```

## Load data
File is loaded from inst/ext where external data files are stored
```{r}
# Load data
load(paste0(here::here(), '/inst/extdata/russian_electoral_interventions_1991-2017.RData'))
rei <- x

incident_count <- nrow(rei)
```

There are `r incident_count` DCID incidents in which Russia was an actor.

# Summary statistics
## Choose variables
Identify variables of interest regarding the means employed, intensity of crises, etc. These variables are all recoded, but their raw values for those that are not obvious are as follows:


```{r}
names(rei)

# Rename variables for readability
rei <- rei %>% dplyr::rename(
  "Year" = year,
  "Target" = target_state,
  "Target COW code" = target_ccode,
  "Target Event" = target_event,
  "Start Date" = event_sdate,
  "End Date" = event_edate,
  "Pro-incumbent" = pro_incumbent,
  "Disinformation campaign" = disinfo,
  "Coup attempt" = coup,
  "Cyberattack" = cyberattack,
  "Material support" = mat_support,
  "Outcome" = outcome,
  "Favorable outcome" = outcome_favorable,
  "Evidence of impact" = rus_impact
)

# Recode some variables to be more intuitive
rei$`Target Event` <- as.character(rei$`Target Event`)
rei$`Target Event` <- plyr::revalue(rei$`Target Event`, c(
  "1" = "National Election",
  "2" = "Referendum",
  "3" = "Political Party Operations"
))

rei$Outcome <- as.character(rei$Outcome)
rei$Outcome <- plyr::revalue(rei$Outcome, c(
  "1" = "Incumbent victory",
  "2" = "Incumbent loss",
  "3" = "Referendum failure",
  "4" = "Referendum success"
))
```

## Summary
Make a table summarizing the data on Russian crisis participation
```{r}
formattable::formattable(rei,
  align = c("c", "l", "l", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c", "c"),
  list(
    "Pro-incumbent" = formattable::color_tile("transparent", "red"),
    "Disinformation campaign" = formattable::color_tile("transparent", "red"),
    "Coup attempt" = formattable::color_tile("transparent", "red"),
    "Cyberattack" = formattable::color_tile("transparent", "red"),
    "Material support" = formattable::color_tile("transparent", "red"),
    "Favorable outcome" = formattable::color_tile("transparent", "lightblue"),
    "Evidence of impact" = formattable::color_tile("transparent", "lightblue")
  )
)
```

## Composite severity variable
There is no variable that directly codes the severity of Russian interference to allow for inter-case comparison.

## Plots
Given the number of cases, summarize them in plot format.
```{r}
DataExplorer::plot_bar(rei)
```

# Save final dataframe
```{r}
saveRDS(rei, paste0(here::here(), "/data/","grayzone_rei.rds"))
```
