---
title: "07_FullData"
author: J Andres Gannon, Erik Gartzke, and Jon Lindsay, Center for Peace and Security
  Studies (cPASS)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  html_notebook:
    fig_height: 8
    fig_width: 12
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document compares case and variable coverage across all the Russia cyber data investigated.

The data comes from the following sources:

ICB comes from Brecher, Michael and Jonathan Wilkenfeld (2000). A Study of Crisis. Ann Arbor: University of Michigan Press.

DCID comes from Valeriano, B., & Maness, R. C. (2014). The dynamics of cyber conflict between rival antagonists, 2001–11. Journal of Peace Research, 51(3), 347–360. https://doi.org/10.1177/0022343313518940

REI comes from Casey, Adam; Way, Lucan Ahmad, 2017, "Russian Electoral Interventions, 1991-2017", https://doi.org/10.5683/SP/BYRQQS, Scholars Portal Dataverse

# Preparation
## Load packages
Pipe operators have trouble loading for individual commands
```{r}
library(magrittr)
library(ggplot2)
```

## Load data
Load the cleaned and subset versions of each dataset
```{r}
# Load data
cpass <- readRDS(paste0(here::here(), '/data/grayzone_cpass.rds'))
icb <- readRDS(paste0(here::here(), '/data/grayzone_icb.rds'))
dcid <- readRDS(paste0(here::here(), '/data/grayzone_dcid.rds'))
rei <- readRDS(paste0(here::here(), '/data/grayzone_rei.rds'))
```

# Compare coverage
Before merging the data, compare the different cases covered by each one

## Prep for coverage comparison
```{r}
# Create identifier variables for each dataset
cpass$cpass <- 1
icb$icb <- 1
dcid$dcid <- 1
rei$rei <- 1

# Ensure proper unit of analysis
dplyr::glimpse(cpass)
dplyr::glimpse(icb)
dplyr::glimpse(dcid)
dplyr::glimpse(rei)

# Fix dcid dates to have a year column
dcid <- tidyr::separate(dcid, 'Start date', c("Start Year", "Start Month", "Start Day"), sep = "-")
dcid <- tidyr::separate(dcid, 'End date', c("End Year", "End Month", "End Day"), sep = "-")

# Rename merging variables to make things easier
cpass <- cpass %>% dplyr::rename('Target' = target,
                                 'Year' = year_start)

icb <- icb %>% dplyr::rename('Target' = Target,
                             'Year' = Year)

dcid <- dcid %>% dplyr::rename('Target' = Target,
                               'Year' = 'Start Year')

rei <- rei %>% dplyr::rename('Target' = Target,
                             'Year' = Year)

# Fix date types
cpass$Year <- as.numeric(cpass$Year)
icb$Year <- as.numeric(icb$Year)
dcid$Year <- as.numeric(dcid$Year)
rei$Year <- as.numeric(rei$Year)
```

## Compare case coverage across datasets
```{r fig.path = '../paper/figures/', dev = c('png')}
# Create dataframe versions of just variables to merge for the match
cpass_coverage <- cpass[ , c('Target', 'Year', 'cpass')]
icb_coverage <- icb[ , c('Target', 'Year', 'icb')]
dcid_coverage <- dcid[ , c('Target', 'Year', 'dcid')]
rei_coverage <- rei[ , c('Target', 'Year', 'rei')]

# Merge datasets with just ID vars
coverage <- dplyr::full_join(cpass_coverage, icb_coverage) %>% dplyr::full_join(., dcid_coverage) %>% dplyr::full_join(., rei_coverage)

# Delete duplicates
coverage <- unique(coverage)

# Convert NAs to 0 to make table easier to read
coverage[is.na(coverage)] <- 0

# Sort chronologically
coverage <- coverage[order(coverage$Year, decreasing = FALSE), ]

# Make html table
formattable::formattable(coverage,
                         align = c("c", "c", "c", "c", "c", "c"),
                         list('cpass' = formattable::color_tile("transparent", "lightblue"),
                              'icb' = formattable::color_tile("transparent", "lightblue"),
                              'dcid' = formattable::color_tile("transparent", "lightblue"),
                              'rei' = formattable::color_tile("transparent", "lightblue")))

# Make Venn diagram
coverage_venn <- coverage[ , c('cpass', 'icb', 'dcid', 'rei')]
gplots::venn(coverage_venn)

```

# Merge data

## Duplicates
First we identify the variables of interest in each dataset and those which are duplicates.
```{r}
# Add ICB merge
icb <- icb %>% dplyr::rename("icb_viol_crisis_cent" = "Crisis Violence Centrality",
                                  "icb_viol_crisis_sev" = "Crisis Violence Severity",
                                  "icb_viol_actor_sev" = "Actor Violence Severity")
icb_join <- icb[ , c('Year', 'Target', 'icb_viol_crisis_cent', 'icb_viol_crisis_sev', 'icb_viol_actor_sev', 'icb')]

# Rename the dcid severity variable so we know the source
dcid <- dcid %>% dplyr::rename("dcid_severity" = "Incident severity")
dcid_join <- dcid[ , c('Target', 'Year', 'dcid', 'Incident', 'dcid_severity')]

# Rename the rei variables so we know the source
rei <- rei %>% dplyr::rename(
  "rei_infoops" = "Disinformation campaign",
  "rei_cyberdisrup" = "Cyberattack",
  "rei_matsupp" = "Material support")

rei_join <- rei[ , c('Target', 'Year', 'rei', 'rei_infoops', 'rei_cyberdisrup', 'rei_matsupp')]
```

## Merge
Merge the old datasets into a single dataframe so they can uploaded to a googlesheet and new variables can be coded.

```{r}
# Merge datasets with just ID vars
cases_old <- dplyr::full_join(icb_join, dcid_join) %>% dplyr::full_join(., rei_join)

# Delete duplicates
cases_old <- unique(cases_old)

# Sort chronologically
cases_old <- cases_old[order(cases_old$Year, decreasing = FALSE), ]
```

# Save final dataframe
This dataframe should now match the one in the google doc for old_cases. SOme duplicate country-years will be merged assuming all other values are identical. They can now have new variables added for our coding of the severity of Russian actions
```{r}
saveRDS(cases_old, paste0(here::here(), "/data/raw/","grayzone_old.rds"))
write.csv(cases_old, paste0(here::here(), "/data/raw/","grayzone_old.csv"))
```

# Reload dataframe with new variables coded
## Clean for merge
The new variables for intensity have been coded by hand into the previous spreadsheet. That can be reloaded now and nerged with the new cases that were similarly coded to analyze the complete version of the dataset. The variables from the new data need to match up with their equivalents in the old data. We'll change the old data to match it up

```{r}
#df_complete <- googlesheets::gs_title("Russian wars")
#df_new <- df_complete %>% googlesheets::gs_read("new_cases")
#df_old <- df_complete %>% googlesheets::gs_read("old_cases")

df_old <- read.csv(paste0(here::here(), "/data/raw/RussianWars-old_cases.csv"))
df_new <- read.csv(paste0(here::here(), "/data/raw/RussianWars-new_cases.csv"))


# Clean data for a clean merge
colnames(df_new)
colnames(df_old)

# Fix variables names
df_new$new <- 1
df_old <- df_old %>% dplyr::rename("incident" = "Incident",
                                   "target" = "Target",
                                   "year_start" = "Year")
```

## Merge
```{r}
# Merge the old and new cases inot a single dataframe
cases_all <- dplyr::full_join(df_new, df_old)

# Sort by date
cases_all <- cases_all[order(cases_all$year_start, decreasing = FALSE), ]

# Inspect quality
summary(cases_all)

# Drop columns we're not using now
vars_keep <- c("target",
               "year_start",
               "resp_convmil",
               "resp_convmil_gro",
               "resp_convmil_airsea",
               "resp_paramil",
               "resp_cyberdisrup",
               "resp_infoops",
               "source",
               "notes",
               "new",
               "incident",
               "dcid",
               "rei",
               "dcid_severity",
               "rei_infoops",
               "rei_cyberdisrup",
               "rei_matsupp")

cases_all <- cases_all[ , vars_keep]

#Save
saveRDS(cases_all, paste0(here::here(), "/data/","grayzone_aggregate_cpass.rds"))
write.csv(cases_all, paste0(here::here(), "/data/","grayzone_aggregate_cpass.csv"))

```


