---
title: "04_ICB"
author: J Andres Gannon, Erik Gartzke, and Jon Lindsay, Center for Peace and Security
  Studies (cPASS)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  html_notebook:
    fig_height: 8
    fig_width: 12
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document will look at coverage of Russian military interventions from 1991-2015 in the ICB data and identify variables of interest.

The ICB data comes from Brecher, Michael and Jonathan Wilkenfeld (2000). A Study of Crisis. Ann Arbor: University of Michigan Press.

# Preparation
## Load data
The ICB data exists in two forms. ICB1 is system-level data and ICB2 is actor-level data. We'll primarily be using the actor-level data but will load both.
```{r}
# Load system-level data
icb_sys <- read.csv(file = paste0(here::here(), '/inst/extdata/icb1v12.csv'))

# Load actor-level data
icb_actor <- read.csv(file = paste0(here::here(), '/inst/extdata/icb2v12.csv'))

```

## Subset
We are only interests in cases since 1991 where Russia was an actor (primarily as an initiator)
```{r}
# Subset to post-1991 cases
icb_actor_rus <- subset(icb_actor, yrtrig > "1990")

# Subset to cases involving Russia
icb_actor_rus <- subset(icb_actor_rus, actor == "RUS")

crisis_count <- nrow(icb_actor_rus)
crisis_list <- icb_actor_rus$crisname
```

There are `r crisis_count` ICB crises post-1991 in which Russia was a crisis actor. This includes `r crisis_list`.

# Summary statistics
Identify variables of interest regarding the means employed, intensity of crises, casualties, etc. From looking over the codebook that includes the following:

CENVIO = centrality of violence

SEVVIO = intensity of violence

OUTCOM = content of crisis outcome

OUTESR = escalation of reduction of tension

TRGTERRA = elapsed time between perception of trigger and termination

RESTERRA = elapsed time between response and termination

GEOG = geographic location of crisis

CRACTLOC = distance of crisis actor from location of crisis

NOACTR = number of actors involved

VIOL = violence associated with crisis actor
```{r}
# Keep identifier variables
vars_id <- c("crisno", "cracid", "actor", "systrgyr", "crisname")
vars_interest <- c("cenvio", "sevvio", "outcom", "outesr", "trgterra", "resterra", "geog", "cractloc", "noactr", "viol")




```
