---
title: "04_ICB"
author: J Andres Gannon, Erik Gartzke, and Jon Lindsay, Center for Peace and Security
  Studies (cPASS)
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  html_notebook:
    fig_height: 8
    fig_width: 12
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document will look at coverage of Russian military interventions from 1991-2015 in the ICB data and identify variables of interest.

The ICB data comes from Brecher, Michael and Jonathan Wilkenfeld (2000). A Study of Crisis. Ann Arbor: University of Michigan Press.

# Preparation
## Load data
The ICB data exists in two forms. ICB1 is system-level data and ICB2 is actor-level data. We'll primarily be using the actor-level data but will load both.
```{r}
# Load system-level data
icb_sys <- read.csv(file = paste0(here::here(), '/inst/extdata/icb1v12.csv'))

# Load actor-level data
icb_actor <- read.csv(file = paste0(here::here(), '/inst/extdata/icb2v12.csv'))

```

## Subset
We are only interests in cases since 1991 where Russia was an actor (primarily as an initiator)
```{r}
# Subset to post-1991 cases
icb_actor_rus <- subset(icb_actor, yrtrig > "1990")

# Subset to cases involving Russia
icb_actor_rus <- subset(icb_actor_rus, actor == "RUS")

crisis_count <- nrow(icb_actor_rus)
crisis_list <- icb_actor_rus$crisname
```

There are `r crisis_count` ICB crises post-1991 in which Russia was a crisis actor. This includes `r crisis_list`.

# Summary statistics
## Choose variables
Identify variables of interest regarding the means employed, intensity of crises, casualties, etc. From looking over the codebook that includes the following:

CENVIO = centrality of violence
  1-4, least to most
SEVVIO = intensity of violence
  1-4, least to most
OUTCOM = content of crisis outcome
  1-4, victory to defeat
OUTESR = escalation of reduction of tension
  1 escalation, 2 de-escalation
TRGTERRA = elapsed time between perception of trigger and termination (count in days)
RESTERRA = elapsed time between response and termination (count in days)
GEOG = geographic location of crisis
  (9) Central Asia
  (10) West Asia (Turkey, Armenia, Azerbaijan, Georgia and Cyprus)
  (11) East Asia
  (12) South-East Asia
  (13) South Asia
  (15) Middle East (countries from the UN’s West Asia category not listed above)
  (20) West Africa
  (21) North Africa
  (22) East Africa
  (23) Southern Africa
  (24) Central Africa
  (30) Euro-Asia (Russia)
  (31) East Europe
  (32) Central Europe (Germany, East Germany, Austria, Switzerland)
  (33) West Europe
  (34) North Europe
  (35) South Europe
  (41) North America
  (42) Central America (including Caribbean countries)
  (43) South America
  (51) Australasia (including Oceania countries)
CRACTLOC = distance of crisis actor from location of crisis
  1-4, closest to farthest
NOACTR = number of actors involved (count)
VIOL = violence associated with crisis actor
  1-4, least to most
```{r}
# Keep identifier variables
vars_id <- c("crisno", "cracid", "actor", "systrgyr", "crisname")
vars_interest <- c("cenvio", "sevvio", "outcom", "outesr", "trgterra", "resterra", "geog", "cractloc", "noactr", "viol")

vars_keep <- c(vars_id, vars_interest)

icb_actor_rus <- icb_actor_rus[ , vars_keep]
```

## Summary
Make a table summarizing the data on Russian crisis participation
```{r}

```
