---
title: "04_FullData"
author: J Andres Gannon, Erik Gartzke, Jon Lindsay, and Peter Schram, Center for Peace and Security
  Studies (cPASS)
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    code_download: true
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
library(magrittr)
library(ggplot2)
```

# Load data
Load the newly created data with all relevant covariates
```{r}
df <- readRDS(paste0(here::here(), '/data/grayzone_model.rds'))
```

# Data summary
Basic summary statistics of the data and what it includes
```{r}
DT::datatable(df)
funModeling::plot_num(df)

df %>%
  dplyr::select(intensity, capdistkm_us, capdistkm_rus, NATOmem_MEM) %>%
  GGally::ggpairs()
```

# Create data versions
We create 3 versions of the data below; known attacks, conservative subset, and expansive subset, based on the inclusion criteria of the universe of cases. For each subset of the data, we create 2 dataframes, one with known values, and a second with imputed values for missing covariates. The covariates with missing data are noted here and are relevant for inclusion of control variables. In most cases, missing covariates are for later years in the dataset (post-2012 or post-2016). In a few cases, there are countries where certain covariates are not known, like missing GDP data for Kosovo and Luxembourg from the World Bank data.
```{r}
# Identify missingness
naniar::gg_miss_var(df)
ExPanDaR::prepare_missing_values_graph(df, ts_id = "year")

# Created imputed version
aout <- df %>%
  dplyr::select(-c(cabbrev1, ccode1, cname2, cabbrev2, ccode2)) %>%
  Amelia::amelia(m = 10,
                 ts = "year",
                 cs = "cname1",
                 polytime = 1,
                 interecs = TRUE,
                 p2s = 1)

# Plot imputation values for the variables with missing data
plot(aout)

Amelia::overimpute(aout, var = "cinc1")
Amelia::overimpute(aout, var = "cinc2")
Amelia::overimpute(aout, var = "cinc_ratio")
Amelia::overimpute(aout, var = "polity1")
Amelia::overimpute(aout, var = "polity2")
Amelia::overimpute(aout, var = "polity_intaxn")
Amelia::overimpute(aout, var = "gdp2_2010const")
```

## Data 1: Known attacks
This includes only country-years where there was a known Russian attack. As a result, there are no values of 0 for the DV. This sample includes non-European states like the US, Canada, Saudi Arabia, and Syria
```{r}
# No imputed values
df_known <- df %>%
  dplyr::filter(intensity > 0)

# Imputed values

```

### Models
We try the 4 different distance metrics
```{r}
m1 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     capdistkm_us +
                     nuclear1,
                   data = df_known,
                   link = "probit")

m2 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     mindistkm_us +
                     nuclear1,
                   data = df_known,
                   link = "probit")

m3 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     capdistkm_rus +
                     nuclear1,
                   data = df_known,
                   link = "probit")

m4 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     mindistkm_us +
                     nuclear1,
                   data = df_known,
                   link = "probit")

summary(m1)
summary(m2)
summary(m3)
summary(m4)
```

## Data 2: Conservative subset
This includes European countries that could have been targets of an attack per our 3 criteria; 1) MID/ICB history with Russia/Soviet Union from 1945-1993, 2) Former Warsaw Pact or Former Soviet Union (FSU) state, 3) contiguity with Russia
```{r}
# No imputed values
df_conserv <- df %>%
  dplyr::filter(relevant_conserv == 1)

# Imputed values

```

### Models
```{r}
m1 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     capdistkm_us +
                     nuclear1,
                   data = df_conserv,
                   link = "probit")

m2 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     mindistkm_us +
                     nuclear1,
                   data = df_conserv,
                   link = "probit")

m3 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     capdistkm_rus +
                     nuclear1,
                   data = df_conserv,
                   link = "probit")

m4 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     mindistkm_us +
                     nuclear1,
                   data = df_conserv,
                   link = "probit")

summary(m1)
summary(m2)
summary(m3)
summary(m4)
```

## Data 3: Expansive subset
Same as the conservative subset, but also adds a fourth criteria of European states that were targets of a Russian attack from 1994-2018. This adds Spain, Germany, Austria, and Malta
```{r}
# No imputed values
df_expansive <- df %>%
  dplyr::filter(relevant_expansive == 1)

# Imputed values

```

### Models
```{r}
m1 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     capdistkm_us +
                     nuclear1,
                   data = df_expansive,
                   link = "probit")

m2 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     mindistkm_us +
                     nuclear1,
                   data = df_expansive,
                   link = "probit")

m3 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     capdistkm_rus +
                     nuclear1,
                   data = df_expansive,
                   link = "probit")

m4 <- oglmx::oglmx(intensity ~ 
                     NATOmem_MEM +
                     mindistkm_us +
                     nuclear1,
                   data = df_expansive,
                   link = "probit")

summary(m1)
summary(m2)
summary(m3)
summary(m4)
```

## Data 4: Full data
This includes all country-years of all states either in Europe or targets of an attack
```{r}

```

# Cumulative results
## Table
```{r, eval = FALSE}
texreg::texreg(
  list(m1),
  custom.model.names = c(
    "All Controls"
  ),
  custom.coef.names = c(
    "Intercept", "Distance from DC", "Distance from Moscow", "NATO membership", "CINC ratio", "Polity Score", "Civil War", "Nuclear power", "GDP"),
  digits = 3,
  ci.force = TRUE,
  ci.force.level = 0.95,
  file = "../paper/model.tex"
)
```

## Figure
```{r, eval = FALSE}
dw1 <- dotwhisker::dwplot(m1, dot_args = list(size = 2.5),
    whisker_args = list(size = 1.5), 
    color = "black") %>%
    relabel_predictors(c(
       capdistkm_us = "Distance from DC",
       capdistkm_rus = "Distance from Moscow",
       NATOmem_MEM = "NATO membership",
       cinc_ratio = "CINC ratio",
       polity1 = "Polity score",
       civilwar = "Civil war",
       nuclear1 = "Nuclear power",
       gdp1 = "GDP"
        )) +
    geom_vline(xintercept = 0) +
    ggtitle("Coefficient Estimates") +
        theme(plot.title = element_text(face="bold", hjust = -0.0),
        legend.justification=c(0, 0), legend.position=c(0, 0),
        legend.title = element_blank()) +
    scale_colour_discrete(breaks=c(
      "Model 1"
      ),
        labels = c(
        "All Controls")) +
    labs(x = "Intensity of Russian Intervention") + 
    theme_bw() +
    theme(text = element_text(size = 24),
        axis.text.x = element_text(colour="black"), 
        axis.text.y = element_text(colour="black"),
        legend.position = c(0.8, 0.01),
        legend.justification = c(0, 0), 
        legend.background = element_rect(colour="grey80"),
        legend.title = element_blank())
```
