---
title: "Appendix"
author: 
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
  html_notebook:
    fig_height: 8
    fig_width: 12
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This appendix provides supplemental information about the dataset of Russian gray zone campaigns introduced in the accompanying paper "After Deterrence: Explaining Conflict Short of War"

# Case selection
The universe of cases was created by first identifying cases of Russian foreign interventions from 3 prior datasets; ICB, DCID, and REI. Code replicating those findings is provided in the appropriate RMarkdown files. These cases were then supplemented with additional cases of Russian interference the authors were able to identify.

A comparison of what cases were covered in each individual dataset is provided here
```{r, echo = FALSE}
# Read dataset
coverage <- readRDS(paste0(here::here(), '/data/coverage.rds'))

# Create table
formattable::formattable(coverage,
                         align = c("c", "c", "c", "c", "c", "c"),
                         list('new' = formattable::color_tile("transparent", "lightblue"),
                              'icb' = formattable::color_tile("transparent", "lightblue"),
                              'dcid' = formattable::color_tile("transparent", "lightblue"),
                              'rei' = formattable::color_tile("transparent", "lightblue")))

```

