---
title: "05a_Figures_Intensity"
author: J Andres Gannon and Erin Werner
date: "`r Sys.Date()`"
output:
  html_document:
    theme: flatly
    code_download: true
    code_folding: hide
    toc: yes
    number_sections: yes
    toc_float:
      collapsed: yes
  html_notebook:
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document visualizes the temporal relationship between deterrence and the severity of Russian cyber operations.

Research assistance for this rmd provided by Erin Werner.

# Preparation
## Load packages
Pipe operators have trouble loading for individual commands
```{r}
library(magrittr)
library(ggplot2)
```

## Load data
For this figure, we load the aggregated dataset that include the new severity codings for all cases in the 3 prior datasets as well as the new cases we have identified.
```{r}
gac <- readRDS(paste0(here::here(), '/data/grayzone_aggregate_new.rds'))
```

# Plotting
## Avergage severity of Russian attacks
We now plot changes in the severity of Russian attacks over the time period in question. The severity will be coded as the highest severity in any given year.
```{r, fig.width = 7.5, fig.height = 7.5}
#plotting the severity of Russian attacks from 1994-2018.
gac <- gac[c("target","year_start","resp_convmil_gro","resp_convmil_airsea",
            "resp_paramil","resp_cyberdisrup","resp_infoops")]

severity <- c()
for(i in 1:nrow(gac)){
  score <- 0
  temp <- gac[i,]
  if(!is.na(temp[3]) & temp[3]==1){ #resp_convmil_gro=5
    score <- 5
  }
  else if(!is.na(temp[4]) & temp[4]==1){ #resp_convmil_airsea=4
    score <- 4
  }
  else if(!is.na(temp[5]) & temp[5]==1){ #resp_paramil=3
    score <- 3
  }
  else if(!is.na(temp[6]) & temp[6]==1){ #resp_cyber_disrup=2
    score <- 2
  }
  else if(!is.na(temp[7]) & temp[7]==1){ #resp_infoops=1
    score <- 1
  }

  severity <- c(severity,score)
}

gac$score <- severity

years <- 1994:2018
severity_avg_annual <- c()
numint <- c()

for(yr in years){
  temp <- gac[which(gac$year_start==yr),]
  if(nrow(temp)>0){
    avg <- mean(temp$score)
    numint <- c(numint,nrow(temp))
    severity_avg_annual <- c(severity_avg_annual, avg)
  }
  else{
    numint <- c(numint,0)
    severity_avg_annual <- c(severity_avg_annual, 0)
  }
}

df <- cbind(years,severity_avg_annual,numint)
df <- as.data.frame(df)

## Start bar chart
g <- ggplot(df, aes(years, severity_avg_annual)) +
  geom_bar(stat = "identity", width = 1, fill = "gray", color = "black") +
  geom_line(aes(y = numint/4)) +
  labs(title = "Russian Interventions (1994-2018)",
       subtitle = "", x = "Year", y = "Average Intensity") +
  scale_x_continuous(breaks = c(1994:2018)) +
  scale_y_continuous(sec.axis = sec_axis(~ . *4, name = "Number of Interventions"),
                     breaks = c(1:5), labels = c("1" = "1 Info ops",
                                                 "2" = "2 Cyber Disrup.",
                                                 "3" = "3 Paramil.",
                                                 "4" = "4 Mil (Air/sea)",
                                                 "5" = "5 Mil (Gro)")) +
  theme_minimal() +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.ticks.x = element_line(),
        title = element_text(size = 18),
        axis.text.x = element_text(size = 12, 
                                   angle = 90, 
                                   vjust = 0.5,
                                   hjust = 0.9),
        axis.text.y = element_text(size = 12),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18)
        )

g

ggsave(paste0("average_intensity_russian_aggression.png"), g ,"png", path = paste0(here::here(), '/paper/figures/'), limitsize = FALSE)
```

## Maximum intensity
We code each column as the maximum intensity of Russian interfence that happened in that year
```{r}
years <- 1994:2018
severity_max_annual <- c()
numint_max <- c()

for(yr in years){
  temp <- gac[which(gac$year_start==yr),]
  if(nrow(temp)>0){
    max <- max(temp$score)
    numint_max <- c(numint_max,nrow(temp))
    severity_max_annual <- c(severity_max_annual, max)
  }
  else{
    numint_max <- c(numint_max,0)
    severity_max_annual <- c(severity_max_annual, 0)
  }
}

df_max <- cbind(years,severity_max_annual,numint_max)
df_max <- as.data.frame(df_max)


#### bar chart
g <- ggplot(df_max, aes(years, severity_max_annual)) +
  geom_bar(stat = "identity", width = 1, fill = "gray", color = "black") +
  geom_line(aes(y = numint_max/4)) +
  labs(title = "Maximum intensity of Russian aggression (1994-2018)",
       subtitle = "", x = "Year", y = "Maximum Intensity") +
  scale_x_continuous(breaks = c(1994:2018)) +
  scale_y_continuous(sec.axis = sec_axis(~.0:20, name = "Number of Russian Interventions"),
                     breaks = c(1, 2, 3, 4, 5), labels = c("1" = "1 Info ops",
                                                           "2" = "2 Cyber Disrup.",
                                                           "3" = "3 Paramil.",
                                                           "4" = "4 Mil (Air/sea)",
                                                           "5" = "5 Mil (Gro)")) +
  theme_minimal() +
  theme(panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        title = element_text(size = 18),
        axis.ticks.x = element_line(),
        axis.text.x = element_text(size = 18, angle = 90,vjust = 0.5,
                                   hjust = 0.9),
        axis.text.y = element_text(size = 14),
        axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18)
  )

g

ggsave(paste0("maximum_intensity_russian_aggression.jpeg"), g, "jpeg", path = paste0(here::here(), '/paper/figures/'), limitsize = FALSE)
```
